Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software

### 2021-08-31 manuscript/manuscript-speed-accuracy.tex
Further manuscript revisions to appease our Reviewer Overlords (who made excellent suggestions).

%\Abstract{
\begin{abstractbox}
  \begin{abstract} % abstract
    \parttitle{Background:} Computational biology provides widely used and
    powerful software tools for testing and making inferences about
    biological data. In the face of rapidly increasing volumes of
    data, heuristic methods that trade software speed for accuracy may
    be employed.  We are have studied these trade-offs using the
    results of a large number of independent software benchmarks, and
    evaluated whether external factors are indicative of accurate
    software.

    \parttitle{Method:} We have extracted accuracy and speed ranks from
    independent benchmarks of different bioinformatic software tools,
    and evaluated whether the speed, author reputation, journal
    impact, recency and developer efforts are indicative of accuracy.

    \parttitle{Results:} We found that software speed, author reputation,
    journal impact, number of citations and age are all unreliable
    predictors of software accuracy. This is unfortunate because
    citations, author and journal reputation are frequently cited
    reasons for selecting software tools. However, GitHub-derived
    records and high version numbers show that the accurate
    bioinformatic software tools are generally the product of many
    improvements over time, often from multiple developers.

    We also find that the field of bioinformatics has a
    large excess of slow and inaccurate software tools, and this is
    consistent across many sub-disciplines. Meanwhile, there are few
    tools that are middle-of-road in terms of accuracy and speed
    trade-offs. We hypothesise that a form of publication-bias
    influences the publication and development of bioinformatic
    software. In other words, software that is intermediate in terms
    of both speed and accuracy may be difficult to publish - possibly
    due to author, editor and reviewer practices. This leaves an
    unfortunate hole in the literature as the ideal tools may fall
    into this gap. For example, high accuracy tools are not always
    useful if years of CPU time are required, while high speed is not
    useful if the results are also inaccurate.
%}
\end{abstract}

@@

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} {\color{red}many believe that} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
{\color{red}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}

%Other factors that are less quantifiable that influence whether a
%%software tool is selected include: whether the documentation is good,
%user-friendly, word-of-mouth and ``used in a similar analysis''
%\cite{Loman2015-bw}. However, citation metrics may be a
%useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.

With the wide adoption of GitHub ($47\%$ of the
{\color{black}\numTools} tools included in this study could be
linked to a GitHub repository), and consequently quantifiable data on software
development time and intensity indicators, such as the number of contributors to code, number of code
changes and versions is available for these
\cite{ray2014large,Dozmorov:2018,Mangul:2019}.

In the following study, we explore factors that may be indicative of
software accuracy. This, in our opinion, should be one of the prime
reasons for selecting a software tool. We have mined the large and
freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
of computational biology software, and manually extracted accuracy and
speed rankings for {\color{black}\numTools~} unique software tools. For
each tool, we have collected measures that may be predictive
of accuracy, and may be subjectively employed by the research
community as a proxy for software quality. These include relative
speed, relative age, the productivity and impact of the corresponding
authors, journal impact and the number of citations.

@@

\begin{figure*}[htb!]
\includegraphics[width=0.9\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  different features of bioinformatic software tools. Spearman’s rho is used to
  infer correlations between metrics such as citations based metrics, the year and relative age of publication, 
  version number, GitHub derived activity measures, and 
  the mean
  relative speed and accuracy rankings. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with an P-value less than 0.05 (corrected for multiple-testing using the Benjami-Hochberg method) are
  indicated with a `$\vartimes$' symbol. The correlations with accuracy are illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} Violin plots of Spearman's correlations between permuted accuracy ranks and 
    different software tool features. 
    The unpermuted correlations are indicated with a red asterix.
    For each benchmark, 1,000 permuted sets of accuracy and speed ranks were generated, and the ranks were normalised to lie between 0 and 1 (see Methods for details).
  Circled asterixs are significant (empirical P-value $< 0.05$, corrected for multiple-testing using the Benjami-Hochberg method).}
\label{fig:allfactors}
\end{figure*}

@@

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number,
{\color{red}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
The results are presented in Figure~\ref{fig:allfactors}A\&B and S5\&S6. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

Data on the number of updates to software tools from GitHub such as
the number of versions, commits and contributors was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg
corrected P-values = {\color{black}0.029, 0.042, 0.063}),
Figure~\ref{fig:allfactors}B.  The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Supplementary
Figures~S5 and~S6}%\ref{fig:S2}
).  We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $3\times3$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}4}
of 9 bins where there was a significant excess or dearth of tools. For
example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.39,
P-value=$3.5\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively). We find that only the ``fast and inaccurate'' extreme class is at approximately the
expected proportions based upon the permutation test (Figure 2B).

@@

The largest difference between the observed and expected software ranks is the reduction in the number of software tools that are
classed as intermediate in terms of both speed and accuracy based on
permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The middle cell
of Figure 2A and left-most violin plot of Figure 2B highlight this extreme, (Z=-6.38,
P-value=$9.0\times 10^{-11}$).

@@

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, which is unfortunate, as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.
%We also found that slow and inaccurate
%software is generally published earlier than fast and accurate tools
%{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
%comparisons with existing software were not required to publish the slow and
%inaccurate tools.

@@

We propose that the full spectrum of software tool accuracies and
speeds serves a useful purpose to the research community. Like negative
results, if honestly reported this information, illustrates to the research community
that certain approaches are not practical research avenues
\cite{fanelli2012negative}. % Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
The current
practices of publishers, editors, reviewers and authors of software
tools therefore may be depriving our community of tools for building effective
and productive workflows.

@@

\textbf{Literature mining:} We identified an initial list of 10 benchmark
articles that satisfy the ALB-criteria. These were identified based
upon previous knowledge of published articles and were supplemented
with several literature searches (e.g., [``benchmark'' AND ``cputime''] was
used to query both GoogleScholar and Pubmed
\cite{Sayers2010-vm,McEntyre2001-fl}). We used these articles to seed
a machine-learning approach for identifying further candidate articles
and to identify new search terms to include.

@@

Where
$\delta$
was a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
$f_{bg}(word)$ and $f_{tr}(word)$ were the frequencies of a $word$ in
the background and training datasets respectively. Word frequencies
were computed by counting the number of times a word appears in the
pool of titles and abstracts, the counts were normalised by the total
number of words in each set.

Thirdly, we also collected a group of candidate benchmark articles by
mining Pubmed for articles that were likely to be benchmarks of
bioinformatic software, these match the terms: “((bioinformatics)
AND (algorithms OR programs OR software)) AND (accuracy OR assessment
OR benchmark OR comparison OR performance) AND (speed OR
time)”. Further terms used in this search were progressively added as
relevant enriched terms were identified in later iterations. The final
query is given in \textbf{supplementary materials}.

@@

\textbf{Data extraction and processing:} for each article that met the
ALB-criteria and contained data on both the accuracy and speed from
their tests, we extracted ranks for each tool. {\color{red}Until published datasets are made available in consistent, machine-readable formats this step is necessarily a manual process -- ranks were extracted from a mixture of manuscript figures, tables and supplementary materials, each data source is documented in Supplementary Table 1. In addition, a variety of accuracy metrics are reported e.g. ``accuracy'', ``AUROC'', ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'', ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc. Our analysis makes the necessarily pragmatic assumption that highly ranked tools on one accuracy metric will also be highly ranked on other accuracy metrics.}
Many articles contained
multiple benchmarks, in these cases we recorded ranks from each of these, the
provenance of which is stored with the accuracy metric and raw speed
and accuracy rank data for each tool {\color{red}(Supplementary Table 1)}. In line with rank-based
statistics, the cases where tools were tied were resolved by using a
midpoint rank (e.g., if tools ranked 3 and 4 are tied, the rank 3.5 was used)
\cite{Mann1947-re}. Each rank extraction was independently verified by
at least one other co-author to ensure both the provenance of the data
could be established and that the ranks were correct. The ranks for
each benchmark were then normalised to lie between 0 and 1 using the
formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is the
number of tools in the benchmark. For tools that were benchmarked
multiple times with multiple metrics (e.g., BWA was evaluated in 6
different articles
\cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
a mean normalised rank was used to summarise the accuracy and speed performance. 
Or, more formally:

@@

For each benchmark of three or more tools, we extracted the published
accuracy and speed ranks. In order to identify whether there was an
enrichment of certain accuracy and speed pairings we constructed a
permutation test. The individual accuracy and speed ranks were
reassigned to tools in a random fashion and each new accuracy and
speed rank pairing was recorded. For each benchmark this procedure was
repeated 1,000 times. These permuted rankings were normalised and
compared to the real rankings to produce the ‘$\vartimes$’ points in
Figure~\ref{fig:allfactors}B and the heatmap and histograms in
Figure~\ref{fig:speedaccuracy}. The heatmap in
Figure~\ref{fig:speedaccuracy} is based upon Z-scores
($Z=\frac{x-\bar{x}}{s}$). For each cell in a $3\times 3$ grid
a Z-score (and corresponding P-value is computed, either with the `pnorm' distribution function in R (Figure 2A) or empirically (Figure 2B)) is computed to illustrate the abundance or lack of tools in
a cell relative to the permuted data.

@@

>>>>

%\Abstract{
\begin{abstractbox}
  \begin{abstract} % abstract
    \parttitle{Background} Computational biology provides widely used and
    powerful software tools for testing and making inferences about
    biological data. In the face of rapidly increasing volumes of
    data, heuristic methods that trade software speed for accuracy may
    be employed.  We are have studied these trade-offs using the
    results of a large number of independent software benchmarks, and
    evaluated whether external factors are indicative of accurate
    software.

    \parttitle{Method} We have extracted accuracy and speed ranks from
    independent benchmarks of different bioinformatic software tools,
    and evaluated whether the speed, author reputation, journal
    impact, recency and developer efforts are indicative of accuracy.

    \parttitle{Results} We found that software speed, author reputation,
    journal impact, number of citations and age are all unreliable
    predictors of software accuracy. This is unfortunate because
    citations, author and journal reputation are frequently cited
    reasons for selecting software tools. However, GitHub-derived
    records and high version numbers show that the accurate
    bioinformatic software tools are generally the product of many
    improvements over time, often from multiple developers.

    \parttitle{Discussion} We also find that the field of bioinformatics has a
    large excess of slow and inaccurate software tools, and this is
    consistent across many sub-disciplines. Meanwhile, there are few
    tools that are middle-of-road in terms of accuracy and speed
    trade-offs. We hypothesise that a form of publication-bias
    influences the publication and development of bioinformatic
    software. In other words, software that is intermediate in terms
    of both speed and accuracy may be difficult to publish - possibly
    due to author, editor and reviewer practices. This leaves an
    unfortunate hole in the literature as the ideal tools may fall
    into this gap. For example, high accuracy tools are not always
    useful if years of CPU time are required, while high speed is not
    useful if the results are also inaccurate.
%}
\end{abstract}

@@

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} {\color{red}many believe that} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
{\color{red}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}
%Other factors that are less quantifiable that influence whether a
%%software tool is selected include: whether the documentation is good,
%user-friendly, word-of-mouth and ``used in a similar analysis''
%\cite{Loman2015-bw}. However, citation metrics may be a
%useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.
{\color{red}\textbf{6. Effective software versioning:} With the wide adoption of
public version-control systems like GitHub, quantifiable data on
software development time and intensity indicators, such as the number
of contributors to code, number of code changes and versions is
now available \cite{ray2014large,Dozmorov:2018,Mangul:2019}.}

In the following study, we explore factors that may be indicative of
software accuracy. This, in our opinion, should be one of the prime
reasons for selecting a software tool. We have mined the large and
freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
of computational biology software, and manually extracted accuracy and
speed rankings for {\color{black}\numTools~} unique software tools. For
each tool, we have collected measures that may be predictive
of accuracy, and may be subjectively employed by the research
community as a proxy for software quality. These include relative
speed, relative age, the productivity and impact of the corresponding
authors, journal impact, number of citations {\color{red}and GitHub activity}.

@@

\begin{figure*}[htb!]
\includegraphics[width=0.9\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  different features of bioinformatic software tools. Spearman’s rho
  is used to infer correlations between metrics such as citations
  based metrics, the year and relative age of publication, version
  number, GitHub derived activity measures, and the mean relative
  speed and accuracy rankings. Red colours indicate a positive
  correlation, blue colours indicate a negative
  correlation. Correlations with a P-value less than 0.05 (corrected
  for multiple-testing using the Benjamini-Hochberg method) are
  indicated with a `$\vartimes$' symbol. The correlations with
  accuracy are illustrated in more detail in \textbf{B}, the
  relationship between speed and accuracy is shown in more detail in
  \textbf{Figure~\ref{fig:speedaccuracy}}. \textbf{B.} Violin plots of Spearman's
  correlations for permuted accuracy ranks and different software
  features.  The unpermuted correlations are indicated with a red
  asterisk.  For each benchmark, 1,000 permuted sets of accuracy and
  speed ranks were generated, and the ranks were normalised to lie
  between 0 and 1 (see Methods for details). Circled asterisks are
  significant (empirical P-value $< 0.05$, corrected for
  multiple-testing using the Benjamini-Hochberg method).}
\label{fig:allfactors}
\end{figure*}

@@

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number,
{\color{red}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
The results are presented in Figure~\ref{fig:allfactors}A\&B, and Additional file 1: Figures S5\&S6. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

Data on the number of updates to software tools from GitHub such as
the number of versions, commits and contributors was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg
corrected P-values = {\color{black}0.029, 0.042, 0.063}),
Figure~\ref{fig:allfactors}B.  The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Additional file 1: Figures S5\&S6}%\ref{fig:S2}
).  We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $3\times3$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}4}
of 9 bins where there was a significant excess or dearth of tools. For
example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.39,
P-value=$3.5\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively). We find that only the ``fast and inaccurate'' extreme class is at approximately the
expected proportions based upon the permutation test (Figure~\ref{fig:speedaccuracy}B).

@@

The largest difference between the observed and expected software ranks is the reduction in the number of software tools that are
classed as intermediate in terms of both speed and accuracy based on
permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The middle cell
of Figure~\ref{fig:speedaccuracy}A and left-most violin plot of Figure~\ref{fig:speedaccuracy}B highlight this extreme, (Z=-6.38,
P-value=$9.0\times 10^{-11}$).

@@

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, which is unfortunate, as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.
%Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?
     {\color{red} Specifically, software tools published early are
       more likely to appear in high impact journals due to their
       perceived novelty and need. Yet without sustained maintenance
       these may be outperformed by subsequent tools, yet early
       publications still accrue citations from users, and all
       subsequent software publications as tools need to be compared
       in order to publish. Subsequent tools are not perceived to be
       as novel, hence appear in ``lower'' tier journals, despite
       being more reliable. Hence, the ``rich'' early publishers get
       richer in terms of citations. Indeed, citation counts are
       mainly predictive of age (Figure~\ref{fig:allfactors}A).}

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are over-represented at both
high and low levels of accuracy{\color{red}, with older tools enriched in this group} (Figure~\ref{fig:speedaccuracy} {\color{red}and Additional file 1: Figure S7}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.

%Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?

%We also found that slow and inaccurate
%software is generally published earlier than fast and accurate tools
%{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
%comparisons with existing software were not required to publish the slow and
%inaccurate tools.

@@

We propose that the full spectrum of software tool accuracies and
speeds serves a useful purpose to the research community. Like
negative results, if honestly reported this information, illustrates
to the research community that certain approaches are not practical
research avenues
\cite{fanelli2012negative}. %Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
The current novelty-seeking practices of many publishers, editors,
reviewers and authors of software tools therefore may be depriving our
community of tools for building effective and productive workflows.
{\color{red} Indeed, the drive for novelty may be an actively harmful
  criteria for the software development community, just as it is for reliable
  and reproducible research
  \cite{brembs2019reliable}. Novelty-criteria for publication may, in
  addition, discourage continual, incremental improvements in code
  post-publication in favour of splashy new tools that are likely to
  accrue more citations. }

{\color{red} In addition we suggest that further efforts be made to
  encourage continual updates to software tools. To paraphrase some of
  the suggestions of Siepel (2019), these efforts may include more
  secure positions for developers, institutional promotion criteria
  include software maintenance, lower publication barriers for
  significant software updates, encourage further funding for software
  maintenance and improvement - not just new tools
  \cite{siepel2019challenges}. If these issues were recognised by
  research managers, funders and reviewers, then perhaps
  the future bioinformatic software tool landscape will be much improved. }

@@

\textbf{Literature mining:} We identified an initial list of 10 benchmark
articles that satisfy the ALB-criteria. These were identified based
upon previous knowledge of published articles and were supplemented
with several literature searches (e.g., [``benchmark'' AND ``cputime''] was
used to query both GoogleScholar and Pubmed
\cite{Sayers2010-vm,McEntyre2001-fl}). We used these articles to seed
a machine-learning approach for identifying further candidate articles
and to identify new search terms to include. {\color{red}This is outlined in Additional file 1: Figure S1. }

@@

Where
$\delta$
was a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
$f_{bg}(word)$ and $f_{tr}(word)$ were the frequencies of a $word$ in
the background and training datasets respectively. Word frequencies
were computed by counting the number of times a word appears in the
pool of titles and abstracts, the counts were normalised by the total
number of words in each set. {\color{red}Additional file 1: Figure S2 shows exemplar word scores. }

Thirdly, we also collected a group of candidate benchmark articles by
mining Pubmed for articles that were likely to be benchmarks of
bioinformatic software, these match the terms: “((bioinformatics)
AND (algorithms OR programs OR software)) AND (accuracy OR assessment
OR benchmark OR comparison OR performance) AND (speed OR
time)”. Further terms used in this search were progressively added as
relevant enriched terms were identified in later iterations. The final
query is given in {\color{red} \textbf{Additional file 1}}.

@@

\textbf{Data extraction and processing:} for each article that met the
ALB-criteria and contained data on both the accuracy and speed from
their tests, we extracted ranks for each tool. {\color{red}Until published datasets are made available in consistent, machine-readable formats this step is necessarily a manual process -- ranks were extracted from a mixture of manuscript figures, tables and supplementary materials, each data source is documented in {\color{red} Additional file 2:} Table S1. In addition, a variety of accuracy metrics are reported e.g. ``accuracy'', ``AUROC'', ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'', ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc. Our analysis makes the necessarily pragmatic assumption that highly ranked tools on one accuracy metric will also be highly ranked on other accuracy metrics.}
Many articles contained
multiple benchmarks, in these cases we recorded ranks from each of these, the
provenance of which is stored with the accuracy metric and raw speed
and accuracy rank data for each tool {\color{red}(Additional file 2: Table S1)}. In line with rank-based
statistics, the cases where tools were tied were resolved by using a
midpoint rank (e.g., if tools ranked 3 and 4 are tied, the rank 3.5 was used)
\cite{Mann1947-re}. Each rank extraction was independently verified by
at least one other co-author to ensure both the provenance of the data
could be established and that the ranks were correct. The ranks for
each benchmark were then normalised to lie between 0 and 1 using the
formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is the
number of tools in the benchmark. For tools that were benchmarked
multiple times with multiple metrics (e.g., BWA was evaluated in 6
different articles
\cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
a mean normalised rank was used to summarise the accuracy and speed performance. 
Or, more formally:

@@

For each benchmark of three or more tools, we extracted the published
accuracy and speed ranks. In order to identify whether there was an
enrichment of certain accuracy and speed pairings we constructed a
permutation test. The individual accuracy and speed ranks were
reassigned to tools in a random fashion and each new accuracy and
speed rank pairing was recorded. For each benchmark this procedure was
repeated 1,000 times. These permuted rankings were normalised and
compared to the real rankings to produce the ‘$\vartimes$’ points in
Figure~\ref{fig:allfactors}B and the heatmap and histograms in
Figure~\ref{fig:speedaccuracy}. The heatmap in
Figure~\ref{fig:speedaccuracy} is based upon Z-scores
($Z=\frac{x-\bar{x}}{s}$). For each cell in a $3\times 3$ grid
a Z-score (and corresponding P-value is computed, either with the `pnorm' distribution function in R (Figure~\ref{fig:speedaccuracy}A) or empirically (Figure~\ref{fig:speedaccuracy}B)) is computed to illustrate the abundance or lack of tools in
a cell relative to the permuted data.

{\color{red} The distributions for each feature and permuted accuracy or speed ranks are shown in {\color{red} Additional file 1: Figures S3\&S4}. Scatter-plots for each pair of features is shown in {\color{red} Additional file 1: Figure S5}. Plots showing the sample sizes for each tool, and feature are shown in {\color{red} Additional file 1: Figure S8}, illustrates a power analysis to show what effect sizes we are likely to detect for our sample sizes. }

@@

\section*{Additional information}

\subsection*{Additional file 1}

{\color{red}Supplementary Figures S1-S8 and the neutral software benchmark reference list the accuracy and speed data is derived from [1-66].}

\subsection*{Additional file 2}

{\color{red}Supplementary Tables S1-S7.}

### 2021-08-31 manuscript/supplementary.tex
Further manuscript revisions to appease our Reviewer Overlords (who made excellent suggestions).

\caption{ {\bf Top:} The number of benchmarks each software tool has appeared in. {\bf Middle:}
  Plot of the amount missing/known data for each of the software features used in this analysis. 
{\bf Bottom:} a power curve analysis to determine if the sample sizes for each set of features is sufficient to detect significant correlations between our features of interest. For example, for speed vs accuracy $N=\numTools$, which is indicated with a vertical line.    
}
\label{fig:numBenchmarks}
\end{figure}

>>>>

\caption{ {\bf Top:} The number of benchmarks each software tool has appeared in. {\bf Middle:}
  Plot of the amount missing/known data for each of the software features used in this analysis. 
{\bf Bottom:} a power curve analysis shows the sample sizes for each set of features is sufficient to detect correlations $r\ge0.2$ between our features of interest with $power \ge 0.8$. For example, for speed vs accuracy $N=\numTools$, which is indicated with a vertical line.    
}
\label{fig:numBenchmarks}
\end{figure}
